Omics-based Single Step Trait Prediction

### 2017-09-17 paper/manuscript.tex
paper, init cover letter

>>>>

% Cover Letter
\section{Cover Letter}
Dear Managing editors of TAG,

We wish to submit the enclosed manuscript "Omics-based Single Step Prediction"
by Westhues et al.\@ for publication in TAG.

Recent studies in maize inbred lines~\cite{Guo2016}, maize hybrids
\cite{Zenke-Philippi2017,Westhues2017}, rice hybrids~\cite{Dan2016,Xu2016} and
humans~\cite{Vazquez2016} have suggested that `omics' predictors, such as the
metabolome and the transcriptome, can improve the prediction of traits with
complex genetic architectures over that achieved with pedigree or genomic data.
Particularly hybrid breeding, where the success of forecasts increases
exponentionally with predictive abilities, could benefit from the use of
predictors that yield more precise estimates of candidate performance.
Currently, the adoption of these novel predictors by breeders might be impeded
by their higher sampling costs compared to pedigree and genomic data.
Here, we present single step prediction of complex traits as a framework for
combining pedigree and/or genomic data with `omics' predictors collected on a
subset of all available genotypes to boost predictive abilities while managing
costs.

Using two extensive phenotypic data sets from a public hybrid silage maize
breeding program and a maize diversity panel, we a) investigate the success of
imputing a quantitative predictor with feature heritabilities less than one and
b) the influence of the fraction of genotypes covered by two predictors on
predictive abilities in single step prediction.

In our current study we obtained the following results:

\begin{enumerate}
\item In both data sets, single step prediction improved predictive abilities
	over those of an individual predictor when the discrepancy between the
	two predictors in terms of predictive ability were moderate to large in
	favor of the incomplete predictor.
\item When 20\%-30\% ($n \approx 25$) of the genotypes were covered by both
	predictors, single step-based predictive abilities converged to those
	obtained with the incomplete predictor.
\end{enumerate}

In conclusion our results demonstrate that single step prediction, using
an established predictor such as pedigree information, can benefit from sampling
a highly performant predictor, such as transcriptomic data, on a subset of all
genotypes.
The framework presented herein should help promoting the use of attractive
`omics' predictors by balancing their advantages with their currently higher
sampling costs.
